Maximum Entropy Spectral Analysis

### 2021-02-08 paper/mesa.tex
Change in paper

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
	\title{Maximum Entropy Spectral Analysis}
	\author{Alessandro \surname{Martini}}
		\email{martini.alessandr@gmail.com}
        \affiliation{Dipartimento di Fisica  Universit\`a di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
        %

	\maketitle

@@

\section{Maximum Entropy Principle} 

@@

\section{Maximum Entropy Spectral Analysis} 

@@

\section{Autoregressive Process Analogy} 

@@

\section{Forecasting}

@@

\section{Analysis of the optimizers}

\section{Example}

\section{Comparison with Welch method}

\section{Appendix 1: MESA solution}

@@

\newpage 
a
\newpage

@@

        \begin{acknowledgments}

          R.~G. acknowledges support from the Deutsche Forschungsgemeinschaft
          (DFG) under Grant No. 406116891 within the Research Training Group
          RTG 2522/1. 
          M.~B. and S.~B. acknowledge support by the EU H2020 under ERC Starting
          Grant, no.~BinGraSp-714626.  
          M.~B.~ acknowledges support from the Deutsche Forschungsgemeinschaft
          (DFG) under Grant No. 406116891 within the Research Training Group
          RTG 2522/1. 
          %

          %
          This research has made use of data, software and/or web tools obtained 
          from the Gravitational Wave Open Science Center (https://www.gw-openscience.org), 
          a service of LIGO Laboratory, the LIGO Scientific Collaboration and the 
          Virgo Collaboration. LIGO is funded by the U.S. National Science Foundation. 
          Virgo is funded by the French Centre National de Recherche Scientifique (CNRS), 
          the Italian Istituto Nazionale della Fisica Nucleare (INFN) and the 
          Dutch Nikhef, with contributions by Polish and Hungarian institutes.
        \end{acknowledgments}

>>>>

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
	\title{Maximum Entropy Spectral Analysis}
	\author{Alessandro \surname{Martini}}
		\email{martini.alessandr@gmail.com}
        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
        %

	\maketitle
	\tableofcontents

@@

\section{Introduction}
\blindtext

\section{Theoretical foundations}

@@

\subsection{Maximum Entropy Spectral Analysis} 

@@

\subsection{Autoregressive Process Analogy} 

@@

\subsection{Forecasting}

\section{Validation of the model}
\subsection{Choosing an autoregressive order}\label{sec:optimizers} 
Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum. If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$. \\ 
In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data. The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$. \\ 
The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact. A compromise between the two behaviour is proposed to be \cite{doi:10.1190/1.1440902}
\begin{equation}\label{eq:MMAx}
M_{max} = 2N / \ln{(2N)}
\end{equation}
Still, this is not a satisfactory answer since it is an upper limit for the estimate of the length $M$.
Various approaches to optimize the order $M$ have been proposed in literature; in the following we present them.
\begin{itemize}
\item \textbf{Final prediction Error (FPE)} 
The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
\begin{equation}
    FPE: \min_{M}E\left((x_t - \hat x_t) ^ 2\right)
\end{equation}
with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.  This criterion is equivalent to the choice of $M$ such that 
\begin{equation}
    \min_{M}P_{m} \frac{N + m + 1}{N - m - 1}
\end{equation}
with $P_M$ defined by the prediction error filter equation. 
\item \textbf{Criterion Autoregressive Transfer function (CAT)}
This second criterion is due to Parzen \cite{bhansali1986}, and requires $M$ to be chosen as the length $m$ the minimizes the quantity 
\begin{equation}
    \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
\end{equation}
It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.

@@

\subsection{Analysis of the optimizers}

\subsection{Comparison with Welch method}

\section{Examples}

\appendix
\section{MESA solution}

@@

%%%%%%%%%%%%%% BELOW THE OLD PART... USEFUL TO TAKE INSPIRATION FROM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
